
# Tips & Tricks {#tips-tricks}

## Debugging

If you've historically used R interactively, you may find it difficult to define functions that get executed at once without your input as Plumber requires. There are a couple of debugging techniques to be aware of when working on your Plumber APIs; these techniques are equally transferrable to debugging your R scripts, packages, or reports.

### Print Debugging

Most programmers first approach debugging by adding print statements to their code in order to inspect the state at some point. In R, `print()` or `cat()` can be used to print out some state. For instance, `cat("i is currently: ", i)` could be inserted in your code to help you ensure that the variable `i` is what it should be at that point in your code.

This approach is equally viable with Plumber. When developing your Plumber API in an interactive environment, this debugging output will be logged to the same terminal where you called `run()` on your API. In a non-interactive production environment, these messages will be included in the API server logs for later inspection.

### Interactive Debugging

Print debugging is an obvious starting point, but most developers eventually wish for something more powerful. In R, this capacity is built in to the `debug()` function. If you're unfamiliar, `debug()` pauses the execution of some function and gives you an interactive session in which you can inspect the current value of internal variables or even proceed through your function one statement at a time.

You can leverage `debug()` when developing your APIs locally by adding a `debug()` call in one of your filters or endpoints and then visiting your API in a client. This offers a powerful technique to use when you want to inspect multiple different variables or interact with the current state of things inside of your function. This is also a good way to get your hands dirty with Plumber and get better acquainted with how things behave at a low level. Consider the following API endpoint:

```r
#' @get /
function(req, res){
  debug()
  
  list(a=123)
}
```

If you run this API locally and then visit the API in a web browser, you'll see your R session switch into debug mode when the request arrives, allowing you to look at the objects contained inside your `req` and `res` objects.

## Testing

// TODO

## Organizing Large Applications

// TOOD

## Swagger

// TODO

## Performance

Run some load tests and show the performance/overhead of running Plumber.

## Cookies

//TODO - see https://plumber.trestletech.com/docs/sessions/ for a good start -- the last section on that page might be good here.

### Setting Unencrypted Cookies {#setting-cookies}

HTTP APIs don't implicitly contain a notion of a "session." Without some additional information, Plumber has no way of ascertaining whether or not two HTTP requests that come in are associated with the same user. Cookies offer a way to commission the client to store some state on your behalf so that selected data can outlive a single HTTP request.

Plumber can both set and receive cookies. For details on setting cookies, see the [Setting Cookies section](#setting-cookies). The API below will send you a random letter, but it remembers your preferences on whether you like capitalized or lower-case letters. 

```{r, echo=FALSE, results='asis'}
include_r("files/apis/06-01-capitalize.R")
```

Since we need to craft a `PUT` request to test this API, we'll use `curl` on the command line to test it. We can start by visiting the `/letter` endpoint and we'll see that the API defaults to a lower-case alphabet. `curl http://localhost:8000/letter`

```{r, echo=FALSE, results='asis'}
pr <- plumber::plumb("files/apis/06-01-capitalize.R")
e <- pr$endpoints[[1]][[2]]
code_chunk(json_serialize(e$exec(req=list(cookies=list()))), "json")
```

If we send a `PUT` request and specify the `capital` parameter, a cookie will be set on the client which will allow the server to accommodate our preference in future requests. In `curl`, you need to specify a file in which you want to save these cookies using the `-c` option. This is a good reminder that clients handle cookies differently -- some won't support them at all -- so be sure that the clients you intend to support with your API play nicely with cookies if you want to use them.

To send a `PUT` request, setting the parameter `capital` to `1`, we could invoke: `curl -c cookies.txt -X PUT --data 'capital=1' "http://localhost:8000/preferences"`. If you print out the `cookies.txt` file, you should now see that it contains a single cookie called `capitalize` with a value of `1`.

We can make another `GET` request to `/letter` to see if it accommodates our preferences. But we'll need to tell `curl` to use the cookies file we just created when sending this request using the `-b` switch: `curl -b cookies.txt http://localhost:8000/letter`. You should now see that the API is returning a random capitalized letter.

This is a simple example showing how to persist user preferences across HTTP requests using cookies. But be aware that the client has the ability to modify or fabricate the cookies that they send to your API. So storing preferences that the user themselves provided in a cookie is not a concern. Storing something with security implications like the level of permissions this client has on your API, however, would be; a malicious user would just need to modify the role saved in their cookie in order to trick your API into giving them more permissions than it should.

There are two common work-arounds to this concern. You can simply store a long (cryptographically) random identifier in the user's cookie, and have some mapping on the server that allows you to lookup the session associated with that random ID. Alternatively, you could use signed/encrypted cookies, as detailed [below](#encrypted-cookies).

### Setting Encrypted Cookies {#encrypted-cookies}

//TODO: See https://plumber.trestletech.com/docs/sessions/ -- it's got a pretty good start
